---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# xmart4

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/caldwellst/xmart4.svg?branch=master)](https://travis-ci.com/caldwellst/xmart4)
<!-- badges: end -->

The goal of xmart4 is to provide easy access to the World Health Organization's
xMart4 API by managing client access tokens for the user and providing simple
functions to view mart and table contents.

## Installation

You can install xmart4 from [GitHub](https://github.com/) with:

```  r
remotes::install_github("caldwellst/xmart4", build_vignettes = TRUE)
```

## Initial setup

```{r child='vignettes/token_setup.rmd'}
```

## Getting data

Once you have sorted out the client tokens, you can start using the simple functions
available in the package to access xMart4 marts and tables.

* `xmart4_mart()` provides a character vector of all available tables and views in a specified mart.
* `xmart4_table()` retrieves data from a specified mart or table.

Using both should just require you to specify mart name, server (UAT or PROD),
and table name (if applicable). It's as easy as opening an R session and going:

```{r example_mart}
library(xmart4)

head(xmart4_mart("GPW13"))
```
Let's access the CONVERT table.
```{r example_convert}
xmart4_table(mart = "GPW13", table = "CONVERT", xmart_server = "UAT")
```

And I can request the top n rows of a table and even supply OData filters
with my request. This is especially useful in instances where tables or views
have many rows and requests may take a long time, so you can explore the table
on a small subset and find useful OData queries to reduce the size of the data
requested.
```{r example_top}
xmart4_table(mart = "GPW13",
             table = "CONVERT",
             top = 2,
             query = "$filter=Input eq 'A'",
             xmart_server = "PROD")
```

Note above, I seamlessly moved between consuming data from PROD and UAT servers.

## Memoisation

`xmart4_api()`, the function underlying requests to the xMart4 API, has cached
functionality based on `memoise::memeoise()` so that calls to the
`xmart4_api()` function are cached in your local memory in a single session.
This means that once you've made a call to access an xMart mart or table/view,
running an identical request will use the cached data rather than re-request the
data from the xMart4 API. This provides large advantages when working with big
tables and views, as the API requests grow quite time consuming as the number of
rows grows. Load the xmart4 package explicitly through `library(xmart4)`, rather
than simply callin functions via `xmart4::fun()`.

This could be problematic if using the xmart4 package to test or consume
data from marts that are being updated throughout an R session. If you need to
ensure that the xmart4 package is making new requests to the API each time, then
you will need to run `memoise::forget(xmart4:::xmart4_api)` to clear the cache
prior to repeating a call. See the documentation of the
[memoise package](https://github.com/r-lib/memoise) for more details.

## Contributions

Additional feature requests should be made through the
[GitHub issues page](https://github.com/caldwellst/xmart4/issues). Any contributions
appreciated and encouraged, but please create an open issue first for discussion
before proceeding with a pull request.
