---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# xmart4

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/caldwellst/xmart4.svg?branch=master)](https://travis-ci.com/caldwellst/xmart4)
<!-- badges: end -->

The goal of xmart4 is to provide easy access to the World Health Organization's
xMart4 API by managing client access tokens for the user and providing simple
functions to view mart and table contents.

## Installation

You can install xmart4 from [GitHub](https://github.com/) with:

```  r
remotes::install_github("caldwellst/xmart4", build_vignettes = TRUE)
```

## Initial setup

```{r child='vignettes/token_setup.rmd'}
```

## Getting data

Once you have sorted out the client tokens, you can start using the simple functions
available in the package to access xMart4 marts and tables.

* `xmart4_mart()` provides a character vector of all available tables and views in a specified mart.
* `xmart4_table()` retrieves data from a specified mart or table.

Using both should just require you to specify mart name, server (UAT or PROD),
and table name (if applicable).

## Contributions

Additional feature requests should be made through the
[GitHub issues page](https://github.com/caldwellst/xmart4/issues). Any contributions
appreciated and encouraged, but please create an open issue first for discussion
before proceeding with a pull request.
